NanoMaLy

### 2015-11-19 paper/pldi16-submission/interactive.tex
down to 11pp

Next, we show how to build an visual interactive debugger
from the traced semantics, by describing the visualization
\emph{state} \ie\ what the user sees at any given moment,
the set of \emph{commands} available to user, and what
they do, and finally how we use a command to \emph{update}
the visualization state. In the sequel, for clarity of
exposition, we assume we have a (global) trace:
$\stepg{e}{\emptysu}{\bullet}{e'}{\vsu'}{\tr}$.

\paragraph{State}
A \emph{visualization state} is ...
The \emph{initial state} is the graph containing the single edge
${e \hookrightarrow \stuck}$.
+ Visualization = ExprGraph + invariants (unique predecessor) -- link to figure.

\paragraph{Commands}
\begin{itemize}
\item \stepforwardsym\ takes a single step forward.
\item \stepbackwardsym\ takes a single step backward.
\item \jumpforwardsym\ takes a ``big'' step forward to the next function call.
\item \jumpbackwardsym\ takes a ``big'' step backward to the previous function call.
\item \stepintosym\ steps into a function call in a sub-term, isolating it from the context.
\item \stepoversym\ steps over a function call in a sub-term.
\end{itemize}

\paragraph{Update}
Figure~\ref{fig:traversing-graph} shows how we update the state for each
command...
%
find   :: Vis -> Cmd -> Expr
update :: Vis -> Cmd -> Vis
%
The initial path $p$ is required for the backward-steps as a node may
have multiple incoming $\leadsto$ edges, \eg
\singlestep{\eplus{1}{2}}{3} and \singlestep{\eplus{2}{1}}{3}.
%
Note that the sub-term edges $\searrow$ allow us to decompose an expression into a
sub-expression and the surrounding context, thus enabling the \stepintosym
and \stepoversym traversals.

data Sel = Expr

data Vis = { graph :: TraceGraph, edges = ExprGraph  }

path :: Vis -> Expr -> [Expr]
stepX :: Sel -> Vis -> Vis(edited)

path v e = graphPath (graph v,  root (v, e),  e)

putBefore, putAfter :: EGraph -> Expr -> Expr -> EGraph
putRoot  :: EGraph -> Expr -> EGraph
getRoot  :: EGraph -> Expr -> Expr

pathToRoot  ::

\begin{figure*}[t]
\centering
\[
\begin{array}{lcl}
\stepforward{G}{p}{e_i}  &\defeq& \left\{\begin{array}{ll}
    e_j, & \text{where } \singlestep{e_i}{e_j} \in G
                         \end{array}\right\} \\ \\
\stepbackward{G}{p}{e_i}  &\defeq& \left\{\begin{array}{ll}
    e_j, & \text{where } \singlestep{e_j}{e_i} \in G \text{ and } e_j \in p
                         \end{array}\right\} \\ \\
\jumpforward{G}{p}{e_i} &\defeq& \text{let } e_j = \stepforward{G}{p}{e_i} \text{ in }
                         \left\{\begin{array}{ll}
                         e_j, & \text{if } e_j = \eapp{v_1}{v_2} \\
                         \jumpforward{G}{p}{e_{j}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\jumpbackward{G}{p}{e_i} &\defeq& \text{let } e_j = \stepbackward{G}{p}{e_i} \text{ in }
                         \left\{\begin{array}{ll}
                         e_j, & \text{if } e_j = \eapp{v_1}{v_2} \\
                         \jumpbackward{G}{p}{e_{j}}, & \text{otherwise}
                         \end{array}\right\} \\ \\
\stepinto{G}{p}{e_i} &\defeq& \left\{\begin{array}{ll}
                         e\sub{x}{v_2}, & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \singlestep{\eapp{v_1}{v_2}}{e\sub{x}{v_2}}
                         \end{array}\right\} \\ \\
\stepover{G}{p}{e_i} &\defeq& \left\{\begin{array}{ll}
                         C[v], & \text{if } e_i = C[\eapp{v_1}{v_2}] \text{ and } \multistep{\eapp{v_1}{v_2}}{v}
                         \end{array}\right\}
\end{array}
\]
\caption{Rules for traversing the reduction graph given a path and
  node. \stepintosym and \stepoversym require a traversal of the
  sub-term edges to decompose $e_i$ into the target expression
  \eapp{v_1}{v_2} and the context $C$.  \ES{these rules are quite ugly and waste space..} }
\label{fig:traversing-graph}
\end{figure*}

>>>>

Next, we show how to build an visual interactive debugger
from the traced semantics, by describing the visualization
\emph{state} \ie\ what the user sees at any given moment,
the set of \emph{commands} available to user, and what
they do, and finally how we use a command to \emph{update}
the visualization state. In the sequel, for clarity of
exposition, we assume we have a (global) trace:
$\stepg{e_0}{\emptysu}{\bullet}{e_n}{\_}{\tr}$, where
$e_0$ and $e_n$ are the \emph{initial} and \emph{final}
expressions respectively.

\paragraph{Visualization State}
%
A \emph{visualization state} $\vstate$ is a \emph{directed graph}
whose vertices are expressions and whose edges are such
that each vertex has at most one predecessor and at most one
successor. In other words, the visualization state looks
like a set of linear lists of expressions as shown in Figure~\ref{fig:FIXTHIS}.
%
The \emph{initial state} is the graph containing a single
edge linking the initial and final expressions.

\paragraph{Commands}
Our debugger supports the following \emph{commands}, each of which
is parameterized by a single expression (vertex) selected from the
(current) visualization state:
%
\begin{itemize}
%
\item \stepforwardsym, \stepbackwardsym:
      show result of a single step forward or backward respectively,
%
\item \jumpforwardsym:
      show result of taking multiple steps (a \emph{``big''} step)
      upto the first beta-reduction forward or backward respectively,
%
\item \stepintosym:
      show result of stepping into a function call in a sub-term,
      isolating it from the context,

\item \stepoversym:
      show result of skipping over a function call in a sub-term.
\end{itemize}

\paragraph{Update}
Figure~\ref{fig:traversing-graph} shows how we update the state
for each command.
%
The procedure $\findExpr{\vstate}{\cmd}{e}$ traverses the $\tr$
and the current visualization state to compute the new
expression that should be added to the visualization state
and $\updState{\vstate}{\cmd}{e}$ then updates the graph
by inserting the new expression appropriately, using one of
the following functions.
%
$\putBefore{\vstate}{e}{e'}$ (resp. $\putAfter{\vstate}{e}{e'}$)
returns the modified version of $\vstate$ where $e$ is the
immediate predecessor of $e'$ (resp. $e'$ the immediate successor of $e$);
%
$\putRoot{\vstate}{e}$ returns the modified version of $\vstate$ extended
with a new vertex $e$ that has no predecessor or successor, and,
%
$\getRoot{\vstate}{e'}$ be the vertex (expression) obtained
by transitively following the predecessors of $e'$ until a
source vertex.
%
Note that the sub-term edges $\searrow$ allow us to decompose an
expression into a sub-expression and the surrounding context, thus
enabling the \stepintosym\ and \stepoversym\ traversals.

\RJ{EXAMPLE of interaction from overview goes here}

%
%
%The initial path $p$ is required for the backward-steps as a node may
%have multiple incoming $\leadsto$ edges, \eg
%\singlestep{\eplus{1}{2}}{3} and \singlestep{\eplus{2}{1}}{3}.
%
